---
title: "Regexp/Date"
output: html_document
---
#grepl
```{r}
text<-"Hi there, do you know who you are voting for?"
grepl("voting",text)

v<-c("a","b","c","d","d")
grepl("b",v)
```

#grep
```{r}
grep("b",v)
grep("d",v)


```

#DATE
```{r}
Sys.Date()
today<-Sys.Date()
class(today)
c<-"1990-01-01"
c<-as.Date(c)
class(c)

mydate<-as.Date("Nov-03-90",format="%b-%d-%y")
# %d day of month
# %m month decimal
# %b month abbrev
# %B month full
# %y 2 digit
# %Y 4 digit

# convert date
df <- data.frame(date = datetxt,
  year = as.numeric(format(datetxt, format = "%Y")),
  month = as.numeric(format(datetxt, format = "%m")),
  day = as.numeric(format(datetxt, format = "%d")),
  hour= as.numeric(format(datetxt, format = "%H")),
  min = as.numeric(format(datetxt, format = "%M")),
  sec = as.numeric(format(datetxt, format = "%S"))
  )

da$START<-as.POSIXct(strptime(data$Trip.Completed.At,format="%Y-%m-%d %H:%M:%S"))

da$COMP<-as.POSIXct(strptime(data$Trip.Started.At,format="%Y-%m-%d %H:%M:%S"))

da$timediff<-difftime(da$START,da$COMP,units="secs")
#"auto", "secs", "mins", "hours", "days", "weeks"



```
#POSIXct convert strings to time format
convert time stamp informaiton, helps you fill out
#strptime, same as posixct
usually use this
date-time conversion fucntions to and from character
```{r}
help(strptime)
#strptime(x, "%m/%d/%y %H:%M:%S")
```


arrange is filter with order

arrange(flights,year,month,day,air_time)
rename(flights,airline_carrier=carrier)
distinct(select(flights,carriers))
mutate #calculation
transmute #only new colum 
sample_n() 
sample_frac(flights,0.1) #get 10% back



